Synthesis of Statically Analyzable Accelerator Networks From Sequential Programs

### 2016-04-25 paper.tex
after john

\numberofauthors{0} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\begin{comment}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Shaoyi Cheng \\%\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{1932 Wallamaloo Lane}\\
       %\affaddr{Wallamaloo, New Zealand}\\
       \email{sh\_cheng@berkeley.edu}
% 2nd. author
\alignauthor
John Wawrzynek \\%\titlenote{The secretary disavows any knowledge of this author's actions.}\\ 
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{P.O. Box 1212}\\
       %\affaddr{Dublin, Ohio 43017-6221}\\
       \email{johnw@eecs.berkeley.edu}
% 3rd. author
}
\end{comment}
\maketitle
\begin{abstract}
This paper describes a general framework in
transforming a sequential program into a network of processes, which are
converted to hardware accelerators through high level synthesis.
A complementing technique is also proposed to perform static deadlock analysis for the generated accelerator network.
The interactions between the accelerators' schedules, the capacity of communication
channels in the network and the memory access mechanisms are all incorporated
into our model, such that potential artificial deadlocks can be detected and resolved \textit{a priori}.

An algorithm optimized for FPGA implementation is then developed and applied through our transformation framework. A set of irregular computation kernels are converted into networks of accelerators on FPGA. 
Compare to hardware accelerators generated without going through our transformation, the
accelerator networks achieve significantly better performance.

@@

\section{Introduction}
As the growth of processor clock frequency came to a stop in the middle of the last decade, the performance boost provided by the continuous advancement of hardware technology for sequential programs could no longer be sustained. The benefits of user transparent CPU architecture optimizations were also diminishing. With the recent advent of FPGA SoCs~\cite{chips:zynq}, offloading computation to hardware accelerators becomes a promising alternative for
achieving better system efficiency. However,
%To create good implementation in the spatial computing
%paradigm however, is far less intuitive compare to imperative programming.
there is a big gap between the inherently parallel mode of operation in hardware accelerators and the intuitive imperative programming paradigm. Even with tool flows like
high level synthesis (HLS), substantial user inputs are still required to create
efficient implementations~\cite{7082747}. To address this problem, we envision a two
step process, where the sequential specification provided by the user is first
transformed to a more parallel representation, from which the conventional HLS
can easily create a high performance compute engine.

\begin{comment}
As parallel computers and more recently, heterogeneous computing platforms become mainstream, it becomes more and more challenging to bridge the gap between the intuitive imperative programming paradigm and the underlying compute substrate.  
In the context of building hardware accelerators on FPGA accelerator
One solution is to leverage more user input, which explicitly specifies the parallelism or

@@

%In this paper, we explore the transformation from sequential programs to networks of processes.
In this paper, we explore this approach by using a modified version of process network (PN), a parallel model of computation, as the target representation for our transformation.
Using FPGA as the computing device, our final implementations are accelerator networks optimized to achieve high throughput in the presence of long data access latency.
Our major contributions include:
\begin{itemize}
    \item We devise a general approach for converting a sequential program to a network
of concurrent processes. Different algorithms can potentially be plugged into the flow, splitting computations between processes in different ways, optimized for the underlying compute substrates.
    \item With FPGA as our target platform, processes are to be converted to hardware accelerators. We
    propose an analysis framework to statically detect and resolve deadlocks in the accelerator networks, ensuring the correctness of the execution. 
    \item We develop an algorithm to generate instances of process networks
particularly amenable to hardware acceleration. Using high level synthesis (HLS), we mapped them onto FPGAs
to demonstrate the benefits of our approach. 
\end{itemize}

@@

The rest of the paper is organized as follows:
Section~\ref{pre} looks at background and related work in process network
%intermediate representation of the sequential program,
and accelerator generation using HLS. Section~\ref{mainConvert} describes the main steps in transforming sequential programs
to networks of processes. When implemented on FPGAs, these networks may experience deadlock issues, which can be statically analyzed and resolved with the method developed in section~\ref{deadlock}. To make our general approach more concrete, we presents a specific algorithm which fits into our PN generation framework in section~\ref{exampleAlgo}.
This algorithm is developed with FPGA as the target platform, thus the generated networks boast great performance when mapped onto the intended 
compute substrate (section~\ref{expEval}). Finally, we conclude the paper in section~\ref{conclude}.
%with discussion and future directions (section~\ref{conclude}).

@@

A characteristic
of KPN is that all data exchange are performed over the FIFOs and the notion of
a shared global memory does not exist in the model. Previous works
converting sequential programs to process networks~\cite{mat2pn}\cite{c2stream}
were able to convert original memory operations to explicit data communication
at the boundary of the generated hardware as the targeted kernels are highly 
regular. In this study however, we target kernels with irregular memory accesses and the final implementations are to be used along side a general purpose processor. We therefore preserve the shared memory, but conceptually model data accesses as inter-process communications.
%the memory model needs to be preserved. It is therefore necessary to conceptually
%translate the behavior of processes accessing memory to that of inter-process
%communication.
%such that our deadlock analysis can also take into account the accelerators' interaction with the memory. 
More specifically, the memory, and its associated interconnect can be seen as a special process which takes in address/data tokens and produces data/acknowledge tokens. In the case where multiple memory ports are exposed, this ``process" does not block on read, as it can choose to serve any port with valid
requests. This behavior is similar to the ``select" function in YAPI~\cite{Kock00yapi:application}, which introduced non-determinism. In our framework however, the non-determinism is addressed by a set of special mechanisms, which will be explained in section~\ref{mainConvert:mem}.

Another necessary deviation from the ideal PN model lies in the capacity of the channels.
In any practical implementations of process network, the FIFOs cannot be unbounded.
Consequently, writes may block and deadlocks can potentially happen. 
These are so-called artificial deadlocks, as their occurrence is introduced by the nonideality of the implementation and the circular dependency formed %in these deadlocks 
always contain operations writing to a full FIFO.
%Note the memory ``process", which does not block on reads,
%can block on writes (responses to other processes), when a large amount of data returned are not consumed promptly. It may therefore also contribute to the development of artificial deadlocks. 
In a process network, % generated from a sequential program, 
these deadlocks, if left unresolved, would cause a mismatch between
the resultant behavior of the original code and the network. To tackle this problem, many approaches were proposed over the years, most of which involves monitoring execution and dynamically incrementing FIFO sizes when deadlock occurs~\cite{parks1995bounded}\cite{Geilen:2003:REK:1765712.1765736}\cite{4217338},
or only targets highly regular kernels.
%This is due to the infeasibility of analyzing deadlock for generic process networks at compile time. 
In our flow, we ensure, by construction, each FIFO is only read/written by a single instruction in each process, which makes the generated PN easier to analyze than a generic PN. We can therefore use a static analysis framework to find the needed FIFO sizes.

\begin{comment}
\subsection{Intermediate Representation for Sequential Programs}
Our PN generation flow is built on top of the LLVM framework~\cite{Lattner:2004:LCF:977395.977673}, which uses a single static assignment (SSA) intermediate representation (IR). 
A typical function in a sequential program is represented with
a control flow graph. This is a two level hierarchical structure. Instructions are contained within basic blocks who are connected by directional edges representing control transfers. 
Meanwhile, every instruction defines a new variable in SSA, thus the dataflow between instructions are explicit and well defined.  
\end{comment}

\subsection{High Level Synthesis of Accelerators}
\label{pre2}
HLS for FPGAs are being active developed in both industry~\cite{tools:vivadohls} and academia~\cite{tools:legup}.
These HLS tools attempt to capture parallelism in a function written with high level languages like C/C++. With the help of various user provided guidance, it generates hardware in the form of RTL. 
%Compute operations and data accesses are allocated and scheduled according to the %dependency constraints extracted from the input program, and the resource %constraints imposed by the target platform. 
The synthesized circuits usually follow a Finite State Machine with Datapath (FSMD) paradigm. Activation of a particular operator in the datapath is associated
with a certain clock cycle and the execution of the computation engine is orchestrated by a synthesized central controller.

@@

\section{From Sequential Program to Process Network}
\label{mainConvert}
Our PN generation flow is built on top of the LLVM framework~\cite{Lattner:2004:LCF:977395.977673}, which uses a single static assignment (SSA) intermediate representation (IR). 
A typical function in a sequential program is represented with
a control flow graph (CFG). This is a two level hierarchical structure. Instructions are contained within basic blocks who are connected by directional edges representing control transfers. 
Meanwhile, every instruction defines a new variable in SSA, thus the data dependencies between instructions are explicit and well defined.  
%The CDFG representing a function is a hierarchical structure.
%Instructions are contained within basic blocks, which are
%connected by directional control edges. 
To generate process
network from this representation, a few main steps are performed:
\begin{itemize}
\begin{comment}
\item \textit{Branch Predication:} The hierarchy in the IR is flattened by branch predication.
The original control transfers are converted to dependency edges 
between branches and other compute/data access operations. 
\end{comment}
\item \textit{Dependency Annotation:} The hierarchy in the IR is flattened by branch predication. Ordering of memory operations is also enforced by the addition of
special dependency edges into the graph.

@@

\item \textit{CFG Reconstruction:} From each set of instruction nodes, a CFG is reconstructed
to form a self-containing process. Dependencies between the instruction
nodes are realized by the communications between the recreated CFGs. 
\begin{comment}
\item \textit{Accelerator Network Generation:} The reconstructed CDFGs are pushed through
high level synthesis, producing an array of communicating hardware accelerators. Linked together
by FIFO channels, these concurrently executing modules form a network which is functionally
equivalent to the original sequential specification.
\end{comment}
\end{itemize}

%\subsection{Branch Predication}
\subsection{Dependency Annotation}
\label{mainConvert:mem}
To facilitate the instruction partitioning, we perform predication to eliminate the basic block boundaries, creating a unified graph of nodes with the
addition of control dependency edges. 
The simplest way to insert these edges is to have every instruction dependent on the last branch of the immediate
predecessors of its container basic blocks. However, we take a more aggressive approach, looking for the earliest control transfer which necessarily leads to the execution of an instruction.
Given an instruction $i$ and its container basic block $bb$, our
flow looks for the nearest set of basic blocks $BB'$ who are not properly post-dominated by $bb$,
and insert the dependency edge between the branch instructions ending each member of $BB'$ and $i$. Compare to the naive predication, this approach use a single control edge to replace a sequence of edges traversing multiple branches, reducing the overall communication
overhead in the network. Also with the elimination of unnecessary edges, the subsequent steps can be more flexible, potentially leading to more efficient final implementation.

Another type of dependencies are implicitly carried by memory accesses. When
two memory accesses target the same location and one of
them is a store, their order in the original program execution
needs to be preserved--a dependency edge is added
between them. We try to avoid adding unnecessary edges as they
can potentially affect the instruction partitioning negatively. 
The flow currently relies on alias analysis to perform partitioning of the memory space. No edges are inserted for accesses to disjoint memory partitions. 
Meanwhile, for code harder to analyze, the flow can take in user annotations, complementing the overly conservative compile time analysis. With the addition of these edges, the ordering of requests sent to the memory
subsystem is constrained to be consistent with the original program execution.
The non-determinism caused by memory selecting amongst multiple requests no longer
impacts the final outcome. In general, for any pair of instructions separated into two sets, the necessary ordering between them is enforced by the cut edges,
which will eventually be converted to communication channels between processes.

@@

The insertion of the sender/receiver of the barrier happens after the instruction partitioning. Depending on if a partition is the sender or receiver of the token,
an extra store/load operation is introduced into the set of instructions, associated
with the basic block where the barrier occurs.
\end{comment}
\subsubsection{Optimizing Communication v.s. Computation}
\label{opt}
 Regardless of exactly how the instructions are partitioned, as each process performs its computation locally, the necessity to communicate with its peers can incur implementation overheads.
For each communication channel, buffer space is to be allocated, and sender/receiver primitives also need to be inserted. Therefore, it is sometimes beneficial to duplicate some nodes after the instruction partitioning. Given sets of instruction nodes created after applying a platform-specific algorithm, for each individual set $G$, the problem of whether to copy other instruction nodes into $G$ can be formulated as an integer programming problem.

@@

\subsection{CFG Reconstruction}
%Need fixing
From each set of instructions generated,
an executable process is created by constructing an independent control flow graph. This step recreates basic blocks to in the new CFG, while inserting communication primitives to send and receive tokens through FIFO channels.
\begin{itemize}
    \item For every instruction $i$  in the current set, recreate
    its container basic block if it's not already in the CFG.
    \item For every instruction $j$ on which $i$ depends, 
    recreate its container basic block if it's not already
    in the CFG.
    \item If $j$ is assigned to another set of instructions, a load operation
    is created locally as a placeholder in its container basic block.
    \item If $i$ is producing operands for instructions assigned to 
    other sets, a store operation is inserted after $i$.
\end{itemize}With 
The set of recreated basic blocks $B$ are thus associated with either the instructions assigned to the current set, or the placeholder instruction supplying them with operands. 
%We then proceed to mirror all the relevant execution paths in the original CFG. We .
%$d$ is used as the entry block for our new CFG. Basic blocks in the path from $d$ to
%every basic block $b \in B$ are added to the CFG 
The insertion of load and store operations is to accommodate various
dependencies between the processes. Note each pair of these inserted communication primitive is associated with a single instruction in the original CFG.The flow of
tokens between them ensures the execution path are synchronized across different CFGs
and the right operands are supplied for computations distributed across the
process network.

@@

\begin{lstlisting}
float foo(float* x, float* prod, int* ind)
{
  float curProd = 1.0;
  for(int i = 0; i < N; i++){
    int curInd = ind[i];
    float curNum = x[curInd];
    curProd = curProd * curNum;
    prod[i] = curProd;
  }
  return curProd;
}
\end{lstlisting}
When this network runs, it writes the same data into the memory, and eventually return
the same value as the original program. As it completes execution, its processes will either return or be blocked on empty FIFO. However, whether it can successfully complete hinges on
the absence of artificial deadlocks, which we are going to look at next.

@@

\section{Artificial Deadlock Analysis}
\label{deadlock}
\subsection{Boundedness of the Generated PNs}
As mentioned in section~\ref{pre1}, it is impossible to have unbounded communication FIFOs between processes in any real PN implementations. For a PN to execute without experiencing artificial deadlocks, it has to be bounded, i.e. certain schedule of the network only requires finite FIFO space.
For a general PN, its boundedness is an undecidable problem~\cite{parks1995bounded}. However, for the process networks generated using our scheme, 
%we can first show it is possible to execute within bounded space, and
%then find the FIFO space necessary to ensure the absence of artificial deadlocks. 
we can easily show their boundedness by constructively create a schedule of operations.

@@

Assume there is an artificial deadlock, we can go around the dependency
cycle and examine the blocked processes $P_b \subseteq PN$. For a process $p_b$
to be blocked at instruction  invocation $j_0$, $j_0$ is either reading from an empty FIFO, or writing to a full FIFO. For the former case, the FIFO is empty because the token's producer instruction $j_0' \in p_b'$ cannot execute, which indicates an earlier instruction  $i_k'$ in $G(p_b')$ is blocked. For the later case, the FIFO is full because the token produced by an earlier invocation $j_{-1}$ has not been taken by its corresponding consumer instruction $j''_{-1} \in p_b''$. This also indicates an earlier instruction $l_n''$ in $G(p'')$ is blocked. Apply this reasoning recursively around the circle of dependencies,
we should have a chain of precedence among instructions invocations $j_0 \succ i_k' \succ ... \succ j_0$
or $j_0 \succ j''_{-1} \succ l_n'' \succ ... \succ j_0$ in $H$. %As all invocations of instructions in the chains come from schedules locally consistent with a single schedule $H$, 
Both chains are self-contradictory and therefore, the scenario for artificial deadlock can never occur.

@@

Unfortunately, in a network implemented in hardware, the FIFO size cannot be easily increased.
%As the producer and consumer instruction instances are rescheduled differently with
%respect to other instructions in their own process, blocking write causes
%artificial deadlock. Increasing the number of slots in the FIFO can help resolve
%this problem. 
To determine the needed buffer space in the channels \textit{a priori}, simulations are sometimes used~\cite{buck1994ptolemy}. It relies 
on using representative dataset as input and% for certain type of problems, simulation results 
may not provide any guarantees.%, especially if the control flow is runtime data dependent. 
In this section, we want to analyze the interaction between FIFO
sizing and intra-process instruction scheduling, 
so we can better understand
their joint effect on the liveness of the process network.

@@

\subsection{Precedence Graph for Deadlock Detection}
To directly detect cycles from the schedules of processes as we have done in 
figure~\ref{fig:2schedules} is unrealistic as each schedule may contain potentially infinite instruction invocations.
However, HLS tools generate a short, repeatable schedule for each process.
We can leverage it to create a more concise representation--precedence graph.

>>>>

\numberofauthors{0} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\begin{comment}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Shaoyi Cheng \\%\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{1932 Wallamaloo Lane}\\
       %\affaddr{Wallamaloo, New Zealand}\\
       \email{sh\_cheng@berkeley.edu}
% 2nd. author
\alignauthor
John Wawrzynek \\%\titlenote{The secretary disavows any knowledge of this author's actions.}\\ 
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{P.O. Box 1212}\\
       %\affaddr{Dublin, Ohio 43017-6221}\\
       \email{johnw@eecs.berkeley.edu}
% 3rd. author
}
\end{comment}
\maketitle
\begin{abstract}
This paper describes a general framework for
transforming a sequential program into a network of processes, which are then
converted to hardware accelerators through high level synthesis.
Also proposed is a complementing technique for performing static deadlock analysis of the generated accelerator network.
The interactions between the accelerators' schedules, the capacity of the communication
channels in the network and the memory access mechanisms are all incorporated
into our model, such that potential artificial deadlocks can be detected and resolved \textit{a priori}.

An algorithm optimized for FPGA implementation is developed and applied through our transformation framework. A set of irregular computation kernels are converted into networks of FPGA accelerators.
Compared to hardware accelerators generated without going through our transformation, the
accelerator networks achieve significantly better performance.

@@

\section{Introduction}
As the growth of processor clock frequency came to a stop in the middle of the last decade, the performance boost provided by the continuous advancement of hardware technology for sequential programs could no longer be sustained. The benefits of user transparent CPU architecture optimizations were also diminishing. With the recent advent of FPGA SoCs~\cite{chips:zynq}, offloading computation to hardware accelerators becomes a promising alternative for
achieving better system efficiency. However,
%To create good implementation in the spatial computing
%paradigm however, is far less intuitive compare to imperative programming.
there is a large gap between the inherently parallel mode of operation in hardware accelerators and the intuitive imperative programming paradigm. Even with tool flows like
high level synthesis (HLS), substantial user inputs are still required to create
efficient implementations~\cite{7082747}. To address this problem, we envision a two
step process, where the sequential specification provided by the user is first
transformed to a more parallel representation, from which conventional HLS
can easily create a high performance compute engine.

\begin{comment}
As parallel computers, and more recently, heterogeneous computing platforms become mainstream, it becomes more and more challenging to bridge the gap between the intuitive imperative programming paradigm and the underlying compute substrate.  
In the context of building hardware accelerators on FPGAs,
one solution is to leverage more user input, which explicitly specifies the parallelism or

@@

%In this paper, we explore the transformation from sequential programs to networks of processes.
In this paper, we explore this approach by using a modified version of the process network (PN), a parallel model of computation, as the target representation for our transformation.
Using FPGA as the computing device, our final implementations are accelerator networks optimized to achieve high throughput in the presence of long data access latency.
Our major contributions include the following.
\begin{itemize}
    \item We devise a general approach for converting a sequential program to a network
of concurrent processes. Different algorithms can potentially be inserted into the flow, splitting computations between processes in different ways, optimized for the underlying compute substrates.
    \item With FPGA as our target platform, processes are to be converted to hardware accelerators. We
    propose an analysis framework to statically detect and resolve deadlocks in the accelerator networks, ensuring the correctness of the execution. 
    \item We develop an algorithm to generate instances of process networks
particularly amenable to hardware acceleration. Using high level synthesis (HLS), we map them onto FPGAs
to demonstrate the benefits of our approach. 
\end{itemize}

@@

The rest of the paper is organized as follows:
Section~\ref{pre} looks at background and related work in process networks
%intermediate representation of the sequential program,
and accelerator generation using HLS. Section~\ref{mainConvert} describes the main steps in transforming sequential programs
to networks of processes. When implemented on FPGAs, these networks may experience deadlock issues, which can be statically analyzed and resolved with the method developed in section~\ref{deadlock}. To make our general approach more concrete, we present a specific algorithm which fits into our PN generation framework in section~\ref{exampleAlgo}.
This algorithm is developed with FPGA as the target platform, thus the generated networks boast great performance when mapped onto the intended 
compute substrate (section~\ref{expEval}). Finally, we conclude the paper in section~\ref{conclude}.
%with discussion and future directions (section~\ref{conclude}).

@@

A characteristic
of KPN is that all data exchange are performed over the FIFOs and the notion of
a shared global memory does not exist in the model. Previous work
converting sequential programs to process networks~\cite{mat2pn}\cite{c2stream}
were able to convert original memory operations to explicit data communication
at the boundary of the generated hardware as the targeted kernels are highly 
regular. In this study however, we target kernels with irregular memory accesses and the final implementations are to be used along side a general purpose processor. We therefore preserve the shared memory, but conceptually model data accesses as inter-process communications.
%the memory model needs to be preserved. It is therefore necessary to conceptually
%translate the behavior of processes accessing memory to that of inter-process
%communication.
%such that our deadlock analysis can also take into account the accelerators' interaction with the memory. 
More specifically, the memory, and its associated interconnect can be seen as a special process which takes in address/data tokens and produces data/acknowledge tokens. In the case where multiple memory ports are exposed, this ``process'' does not block on read, as it can choose to serve any port with valid
requests. This behavior is similar to the ``select'' function in YAPI~\cite{Kock00yapi:application}, which introduced non-determinism. In our framework however, the non-determinism is addressed by a set of special mechanisms, as explained in section~\ref{mainConvert:mem}.

Another necessary deviation from the ideal PN model lies in the capacity of the channels.
In any practical implementations of process networks, the FIFOs cannot be unbounded.
Consequently, in cases with circular dependencies, when buffers fill up writes may block and deadlocks can occur. 
These are so-called artificial deadlocks, as their occurrence is introduced by the nonideality of the implementation. % and the circular dependency formed %in these deadlocks 
always contain operations writing to a full FIFO.
%Note the memory ``process", which does not block on reads,
%can block on writes (responses to other processes), when a large amount of data returned are not consumed promptly. It may therefore also contribute to the development of artificial deadlocks. 
In a process network, % generated from a sequential program, 
these deadlocks, if left unresolved, would cause a mismatch between
the resultant behavior of the original code and the network. To tackle this problem, many approaches were proposed over the years, most of which involves monitoring execution and dynamically incrementing FIFO sizes when deadlock occurs~\cite{parks1995bounded}\cite{Geilen:2003:REK:1765712.1765736}\cite{4217338},
or only targets highly regular kernels.
%This is due to the infeasibility of analyzing deadlock for generic process networks at compile time. 
In our flow, we ensure, by construction, each FIFO is only read/written by a single instruction in each process, which makes the generated PN easier to analyze than a generic PN. We can therefore use a static analysis framework to find the needed FIFO sizes.

\begin{comment}
\subsection{Intermediate Representation for Sequential Programs}
Our PN generation flow is built on top of the LLVM framework~\cite{Lattner:2004:LCF:977395.977673}, which uses a single static assignment (SSA) intermediate representation (IR). 
A typical function in a sequential program is represented with
a control flow graph. This is a two level hierarchical structure. Instructions are contained within basic blocks which are connected by directional edges representing control transfers. 
Meanwhile, every instruction defines a new variable in SSA, thus the dataflow between instructions are explicit and well defined.  
\end{comment}

\subsection{High Level Synthesis of Accelerators}
\label{pre2}
HLS for FPGAs are being actively developed in both industry~\cite{tools:vivadohls} and academia~\cite{tools:legup}.
These HLS tools attempt to capture parallelism in a function written with high level languages like C/C++. With the help of various user provided guidance these tools generates hardware in the form of RTL. 
%Compute operations and data accesses are allocated and scheduled according to the %dependency constraints extracted from the input program, and the resource %constraints imposed by the target platform. 
The synthesized circuits usually follow a Finite State Machine with Datapath (FSMD) paradigm. Activation of a particular operator in the datapath is associated
with a certain clock cycle and the execution of the computation engine is orchestrated by a synthesized central controller.

@@

\section{From Sequential Program to Process Network}
\label{mainConvert}
Our PN generation flow is built on top of the LLVM framework~\cite{Lattner:2004:LCF:977395.977673}, which uses a single static assignment (SSA) intermediate representation (IR). 
A typical function in a sequential program is represented with
a control flow graph (CFG). This is a two level hierarchical structure. Instructions are contained within basic blocks which are connected by directional edges representing control transfers. 
Meanwhile, every instruction defines a new variable in the SSA representation, thus the data dependencies between instructions are explicit and well defined.  
%The CDFG representing a function is a hierarchical structure.
%Instructions are contained within basic blocks, which are
%connected by directional control edges. 
To generate the process
network from this representation, a few main steps are performed:
\begin{itemize}
\begin{comment}
\item \textit{Branch Predication:} The hierarchy in the IR is flattened by branch predication.
The original control transfers are converted to dependency edges 
between branches and other compute/data access operations. 
\end{comment}
\item \textit{Dependency Annotation:} The hierarchy in the IR is flattened by branch predication. Ordering of memory operations is also enforced by the addition of
special dependency edges into the graph.

@@

\item \textit{CFG Reconstruction:} From each set of instruction nodes, a CFG is reconstructed
to form a self-containing process. Dependencies between the instruction
nodes are realized by the communication between the recreated CFGs. 
\begin{comment}
\item \textit{Accelerator Network Generation:} The reconstructed CDFGs are pushed through
high level synthesis, producing an array of communicating hardware accelerators. Linked together
by FIFO channels, these concurrently executing modules form a network which is functionally
equivalent to the original sequential specification.
\end{comment}
\end{itemize}

%\subsection{Branch Predication}
\subsection{Dependency Annotation}
\label{mainConvert:mem}
To facilitate the instruction partitioning, we perform predication to eliminate the basic block boundaries, creating a unified graph of nodes with the
addition of control dependency edges. 
The simplest way to insert these edges is to have every instruction dependent on the last branch of the immediate
predecessors of its container basic blocks. However, we take a more aggressive approach, looking for the earliest control transfer which necessarily leads to the execution of an instruction.
Given an instruction $i$ and its container basic block $bb$, our
flow looks for the nearest set of basic blocks $BB'$ who are not properly post-dominated by $bb$,
and insert the dependency edge between the branch instructions ending each member of $BB'$ and $i$. Compared to naive predication, this approach use a single control edge to replace a sequence of edges traversing multiple branches, reducing the overall communication
overhead in the network. Also with the elimination of unnecessary edges, the subsequent steps can be more flexible, potentially leading to a more efficient final implementation.

Another type of dependency is implicitly carried by memory accesses. When
two memory accesses target the same location and one
is a store, their order in the original program execution
needs to be preserved---a dependency edge is added
between them. We try to avoid adding unnecessary edges as they
can potentially affect the instruction partitioning negatively. 
The flow currently relies on alias analysis to perform partitioning of the memory space. No edges are inserted for accesses to disjoint memory partitions. 
Meanwhile, for code harder to analyze, the flow can take in user annotations, complementing the overly conservative compile time analysis. With the addition of these edges, the ordering of requests sent to the memory
subsystem is constrained to be consistent with the original program execution.
The non-determinism caused by memory selecting among multiple requests no longer
impacts the final outcome. In general, for any pair of instructions separated into two sets, the necessary ordering between them is enforced by the cut edges,
which will eventually be converted to communication channels between processes.

@@

The insertion of the sender/receiver of the barrier happens after the instruction partitioning. Depending on if a partition is the sender or receiver of the token,
an extra store/load operation is introduced into the set of instructions, associated
with the basic block where the barrier occurs.
\end{comment}
\subsubsection{Optimizing Communication vs. Computation}
\label{opt}
 Regardless of exactly how the instructions are partitioned, as each process performs its computation locally, the necessity to communicate with its peers can incur implementation overheads.
For each communication channel, buffer space is to be allocated, and sender/receiver primitives also need to be inserted. Therefore, it is sometimes beneficial to duplicate some nodes after the instruction partitioning. Given sets of instruction nodes created after applying a platform-specific algorithm, for each individual set $G$, the problem of whether to copy other instruction nodes into $G$ can be formulated as an integer programming problem.

@@

\subsection{CFG Reconstruction}
%Need fixing
From each set of instructions generated,
an executable process is created by constructing an independent control flow graph. This step recreates basic blocks to in the new CFG, while inserting communication primitives to send and receive tokens through FIFO channels.
\begin{itemize}
    \item For every instruction $i$  in the current set, recreate
    its container basic block if it's not already in the CFG.
    \item For every instruction $j$ on which $i$ depends, 
    recreate its container basic block if it's not already
    in the CFG.
    \item If $j$ is assigned to another set of instructions, a load operation
    is created locally as a placeholder in its container basic block.
    \item If $i$ is producing operands for instructions assigned to 
    other sets, a store operation is inserted after $i$.
\end{itemize}With 
The set of recreated basic blocks $B$ are thus associated with either the instructions assigned to the current set, or the placeholder instruction supplying them with operands. 
%We then proceed to mirror all the relevant execution paths in the original CFG. We .
%$d$ is used as the entry block for our new CFG. Basic blocks in the path from $d$ to
%every basic block $b \in B$ are added to the CFG 
The insertion of load and store operations is to accommodate various
dependencies between the processes. Note that each pair of these inserted communication primitive is associated with a single instruction in the original CFG. The flow of
tokens between them ensures the execution path are synchronized across different CFGs
and the right operands are supplied for computations distributed across the
process network.

@@

\begin{lstlisting}
float foo(float* x, float* prod, int* ind)
{
  float curProd = 1.0;
  for(int i = 0; i < N; i++){
    int curInd = ind[i];
    float curNum = x[curInd];
    curProd = curProd * curNum;
    prod[i] = curProd;
  }
  return curProd;
}
\end{lstlisting}
When this network runs, it writes the same data into the memory, and eventually return
the same value as the original program. As it completes execution, its processes will either return or be blocked on an empty FIFO. However, whether it can successfully complete hinges on
the absence of artificial deadlocks---which we look at next.

@@

\section{Artificial Deadlock Analysis}
\label{deadlock}
\subsection{Boundedness of the Generated PNs}
As mentioned in section~\ref{pre1}, it is impossible to have unbounded communication FIFOs between processes in any real PN implementations. For a PN to execute without experiencing artificial deadlocks, it has to be bounded, i.e. certain schedule of the network only requires finite FIFO space.
For a general PN, its boundedness is an undecidable problem~\cite{parks1995bounded}. However, for the process networks generated using our scheme, 
%we can first show it is possible to execute within bounded space, and
%then find the FIFO space necessary to ensure the absence of artificial deadlocks. 
we can easily show their boundedness by constructively creating a schedule of operations.

@@

Assuming there is an artificial deadlock, we can go around the dependency
cycle and examine the blocked processes $P_b \subseteq PN$. 
For a process $p_b$ to be blocked at instruction invocation $j_0$, $j_0$ is either reading from an empty FIFO, or writing to a full FIFO. 
For the former case, the FIFO is empty because the token's producer instruction $j_0' \in p_b'$ cannot execute, which indicates an earlier instruction  $i_k'$ in $G(p_b')$ is blocked. 
For the later case, the FIFO is full because the token produced by an earlier invocation $j_{-1}$ has not been taken by its corresponding consumer instruction $j''_{-1} \in p_b''$. 
This also indicates an earlier instruction $l_n''$ in $G(p'')$ is blocked. 
Applying this reasoning recursively around the circle of dependencies,
we should have a chain of precedence among instructions invocations $j_0 \succ i_k' \succ ... \succ j_0$
or $j_0 \succ j''_{-1} \succ l_n'' \succ ... \succ j_0$ in $H$. %As all invocations of instructions in the chains come from schedules locally consistent with a single schedule $H$, 
Both chains are self-contradictory and therefore, the scenario for artificial deadlock can never occur.

@@

Unfortunately, in a network implemented in hardware, the FIFO size cannot be easily increased.
%As the producer and consumer instruction instances are rescheduled differently with
%respect to other instructions in their own process, blocking write causes
%artificial deadlock. Increasing the number of slots in the FIFO can help resolve
%this problem. 
To determine the required buffer space in the channels \textit{a priori}, simulations are sometimes used~\cite{buck1994ptolemy}. It relies 
on using representative datasets as input and% for certain type of problems, simulation results 
may not provide any guarantees.%, especially if the control flow is runtime data dependent. 
In this section, we want to analyze the interaction between FIFO
sizing and intra-process instruction scheduling, 
so we can better understand
their joint effect on the liveness of the process network.

@@

\subsection{Precedence Graph for Deadlock Detection}
To directly detect cycles from the schedules of processes as we have done in 
figure~\ref{fig:2schedules} is unrealistic as each schedule may contain potentially infinite instruction invocations.
However, HLS tools generate a short, repeatable schedule for each process.
We can leverage this fact to create a more concise representation---a precedence graph.

An example of a precedence graph is shown in figure~\ref{fig:predgraph}, the schedules of instruction invocations are condensed into a small graph whose edges carry vector weights. 
%As illustrated in the figure, in a schedule of a multiple level loop nest, the instruction invocations can be annotated with subscripts representing their position in the iteration space. For instance, instruction executed in the second iteration of the inner loop within the third iteration of the outer loop has subscript (3,2). Also, outer loop instruction
%invocations carry 1 at the position of the vector corresponding to the inner loop level. 
The weights of the $DepS$ edges are simply the worst case differences between the vector prefixes of invocations of a pair of instructions. 
Similarly, the weights of $DepF$ edges correspond to the differences in prefixes between
the consumer and producer instruction invocations. Note this is directly related
to the number of slots in the FIFO. For our example, the edge from store(r1Q,r1) to its consumer load(r1Q) carries a weight of (0,-1) when the channel has just one slot.
It means the previous invocation of the consumer instruction needs to complete before the producer instruction can write more tokens. The $DepD$ edges, on the other hand, always have weight $\vec{0}$ as the consumer depends on the producer with the same iteration vector. In essence, this graph
summarizes the necessary precedence between instruction invocations.
